---
title: "Improving memory for and production of singular <i>they</i> pronouns: Experiment 1B"
author: "Bethany Gardner"
date: "03/24/2022"
output: 
  github_document:
    toc: true
    toc_depth: 3
  pdf_document:
    toc: true
    toc_depth: 3
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(magrittr)
library(lmerTest)
library(buildmer)
library(sjmisc)
```

This is a replication of the first experiment. Everything is identical,
except that the production task occurs before the memory task.

# Load Data

Read data, preprocessed from Qualtrics output. See
data/exp1b_data_readme for more details.

```{r}
d <- read.csv("../data/exp1b_data.csv", stringsAsFactors=TRUE)
str(d)
```

Set up contrast coding. The first contrast compares they to he+she. The
second contrast compares he to she.

```{r}
contrasts(d$Pronoun) = cbind("they vs he+she"=c(.33,.33,-.66), 
                             "he vs she"=c(-.5,.5, 0))
contrasts(d$Pronoun)
```

Split data by task, and only keep pronoun questions (not the job or pet
questions) in memory dataframe.

```{r}
m <- d %>% filter(M_Type=="pronoun")
p <- d %>% filter(Task=="production")
```

Combine memory and production trials to make one row for each character.

```{r}
#Get pronoun memory and production observations. Filter out memory for job and pet questions, and introduction pilot task questions.
mp <- d %>% filter(Task != "introduction" & 
                   M_Type != "job" &
                   M_Type != "pet") 

#Just take columns used in model
m_temp <- mp %>% select(M_Acc, Pronoun, Name, SubjID) %>%
  filter(!is.na(M_Acc)) #Take out empty rows that were other question types

#Get production accuracy column
p_temp <- mp %>% select(P_Acc)  %>%
    filter(!is.na(P_Acc)) #Take out empty rows that were other question types

#Combine
mp <- cbind(m_temp, p_temp) 
str(mp)
```

# Memory

### Descriptive Stats

Mean accuracy for all three memory question types.

```{r}
prop.table(table(d$M_Type, d$M_Acc), margin=1)
```

Mean accuracy, split by pronoun type.

```{r}
prop.table(table(m$Pronoun, m$M_Acc), margin=1)
```

96% of participants selected they/them at least once.

```{r}
they_m <- d %>% filter(M_Response=="they/them") %>%
  summarize(n=n_distinct(SubjID)) 

they_m/(n_distinct(d$SubjID))

```

### Model

Start with model that has random intercepts and slopes for participant
and item. Maximal model has by-participant random intercepts only.

```{r}
model_m_full <- M_Acc ~ Pronoun + (1 + Pronoun|SubjID) + (1 + Pronoun|Name)

model_m <- buildmer(model_m_full, data=m, 
                    family='binomial', direction=c('order'))
summary(model_m)
```

Convert to odds:

```{r}
exp(0.94930) #intercept (mean)
exp(1.55205) #they/them vs. he/him + she/her
```

-   The intercept is significant (p\<.001), such that participants are
    2.58 times more likely to answer correctly than incorrectly across
    all pronoun types.

-   The contrast between they/them and he/him + she/her is significant
    (p\<.001), such that participants are 4.72 times more likely to get
    he/him and she/her right than they/them.

-   The contrast between he/him and she/her is not significant.

# Production

### Descriptive Stats

Mean accuracy, split by pronoun type. Accuracy for producing they/them
is lower than accuracy for remembering they/them.

```{r}
prop.table(table(p$Pronoun, p$P_Acc), margin=1)
```

71% of participants produced they/them at least once.

```{r}
they_p <- d %>% filter(P_Pronoun=="they/them") %>%
  summarize(n=n_distinct(SubjID)) 

they_p/(n_distinct(d$SubjID))

```

### Model

Start with model that has random intercepts and slopes for participant
and item, using same specifications as before. Maximal model has random
intercepts by participant and item, and no random slopes.

```{r}
model_p_full <- P_Acc ~ Pronoun + 
  (1 + Pronoun|SubjID) + (1 + Pronoun|Name)

model_p <- buildmer(model_p_full, data=p, 
          family='binomial', direction=c('order'))
summary(model_p)
```

Convert to odds:

```{r}
exp(1.10833) #intercept (mean)
exp(2.476427) #they/them vs. he/him + she/her
```

-   The intercept is significant (p\<.001), such that participants are
    3.02 times more likely to answer correctly than incorrectly across
    all pronoun types.

-   The contrast between they/them and he/him + she/her is significant
    (p\<.001), such that participants are 11.83 times more likely to get
    he/him and she/her right than they/them.

-   The contrast between he/him and she/her is not significant.

# Memory Predicting Production

### Descriptive Stats

Combining the two measures, there are 4 possible patterns: getting both
right, getting both wrong, getting just memory right, and getting just
production right.

```{r}
mp_acc <- mp %>% 
          mutate(BothRight=ifelse(M_Acc==1 & P_Acc==1, 1, 0)) %>%
          mutate(BothWrong=ifelse(M_Acc==0 & P_Acc==0, 1, 0)) %>%
          mutate(MemOnly=ifelse(M_Acc==1 & P_Acc==0, 1, 0)) %>%
          mutate(ProdOnly=ifelse(M_Acc==0 & P_Acc==1, 1, 0)) %>%
          pivot_longer(cols=c(BothRight, BothWrong, MemOnly, ProdOnly),
                       names_to="Combined_Accuracy") %>%
          group_by(Pronoun, Combined_Accuracy) %>%
          summarise(m=mean(value))
mp_acc
```

### Model

Model predicting production accuracy with pronoun type and memory
accuracy. Here, the maximal model includes no random effects.

```{r}
model_mp_full <- P_Acc ~ M_Acc * Pronoun + 
  (1 + Pronoun|SubjID) + (1 + Pronoun|Name)

model_mp <- buildmer(model_mp_full, data=mp, 
            family='binomial', direction=c('order'))
summary(model_mp)
```

Convert to odds:

```{r}
exp(1.49113)  #memory accuracy
exp(-1.15555) #they/them vs. he/him + she/her * memory accuracy
```

-   The effect of memory accuracy is significant (p\<.001), such that
    participants are 4.44x more likely to get the production right if
    they got the memory right.

-   Significant interaction between pronoun type (they/them vs. he/him +
    she/her) and memory accuracy (p\<.05) (odds 0.31). The relative
    difficulty of they/them was attenuated when the participant had
    correctly remembered the character's pronoun during the memory phase
    of the task.
