---
title: "Improving memory for and production of singular <i>they</i> pronouns: Experiment 1"
author: "Bethany Gardner"
date: "03/24/2022, 04/08/2023"
output: 
  pdf_document:
    toc: true
    toc_depth: 3
  github_document:
    toc: true
    toc_depth: 3
editor_options: 
  markdown: 
    wrap: 72
---

```{r}
#| label: exp1-setup
#| include: false

library(tidyverse)
library(magrittr)
library(insight)
library(lme4)
library(lmerTest)
library(buildmer)

options(dplyr.summarise.inform = FALSE)
```

# Load Data

Read data, preprocessed from Qualtrics output. See data/exp1a_data_readme for more details.

```{r}
#| label: exp1-load

exp1a_d_all <- read.csv("data/exp1a_data.csv", stringsAsFactors = TRUE)
str(exp1a_d_all)
```

Set up contrast coding. The first contrast compares they to he+she. The second contrast compares he to she.

```{r}
#| label: exp1-contrasts

contrasts(exp1a_d_all$Pronoun) <- cbind(
  "they vs he+she" = c(.33, .33, -.66),
  "he vs she"      = c(-.5, .5, 0))
contrasts(exp1a_d_all$Pronoun)
```

Combine pronoun memory and production trials to make one row for each character.

```{r}
#| label: exp1-reshape

m_temp <- exp1a_d_all %>%
  filter(M_Type == "pronoun") %>%
  select(SubjID, Name, Pronoun, M_Response, M_Acc)
p_temp <- exp1a_d_all %>%
  filter(!is.na(P_Acc)) %>%
  select(SubjID, Name, Pronoun, P_Pronoun, P_Acc)

exp1a_d <- left_join(m_temp, p_temp, by = c("SubjID", "Name", "Pronoun"))
remove(m_temp, p_temp)

str(exp1a_d)
```

# Memory

## Descriptive Stats

Mean accuracy for all three memory question types.

```{r}
#| label: exp1-memory-all-means

exp1a_d_all %>%
  filter(!is.na(M_Acc)) %>%
  group_by(M_Type) %>%
  summarise(
    mean = mean(M_Acc) %>% round(2),
    sd   = sd(M_Acc)   %>% round(2)
  )
```

Mean accuracy, split by pronoun type.

```{r}
#| label: exp1-memory-pronoun-means

exp1a_d %>%
  mutate(Pronoun_Group = ifelse(Pronoun == "they/them", "They", "He + She")) %>%
  group_by(Pronoun_Group) %>%
  summarise(
    mean = mean(M_Acc) %>% round(2),
    sd   = sd(M_Acc)   %>% round(2)
  )
```

94% of participants selected they/them at least once.

```{r}
#| label: exp1-memory-usedThey

exp1a_r_memory_usedThey <- exp1a_d %>%
  filter(M_Response == "they/them") %>%
  summarise(n = n_distinct(SubjID))

exp1a_r_memory_usedThey / (n_distinct(exp1a_d$SubjID))
```

## Model

Start with model that has random intercepts and slopes for participant and item. Specifying the direction as "order" in buildmer will identify the maximal model that will converge. However, it doesn't continue to backward stepwise elimination. This results in a model with random intercepts and slopes by participant, and random intercepts by item.

```{r}
#| label: exp1-memory-model
#| cache: true

exp1a_m_memory <- buildmer(
  formula = M_Acc ~ Pronoun + (1 + Pronoun | SubjID) + (1 + Pronoun | Name),
  data = exp1a_d, family = binomial,
  buildmerControl(direction = "order")
)

summary(exp1a_m_memory)
```

Convert to odds:

```{r}
#| label: exp1-memory-OR

# intercept (mean)
exp1a_m_memory@model %>% get_intercept() %>% exp()

# they/them vs. he/him + she/her
exp1a_m_memory@model %>%
  get_parameters() %>%
  filter(Parameter == "Pronounthey vs he+she") %>%
  pull(Estimate) %>%
  exp()
```

-   The intercept is significant (p\<.001), such that participants are 2.16 times more likely to answer correctly than incorrectly across all pronoun types.

-   The contrast between they/them and he/him + she/her is significant (p\<.001), such that participants are 5.15 times as likely to get he/him and she/her right than they/them.

-   The contrast between he/him and she/her is not significant.

# Production

## Descriptive Stats

Mean accuracy, split by pronoun type.

```{r}
#| label: exp1-production-pronoun-means

exp1a_d %>%
  mutate(Pronoun_Group = ifelse(Pronoun == "they/them", "They", "He + She")) %>%
  group_by(Pronoun_Group) %>%
  summarise(
    mean = mean(P_Acc) %>% round(2),
    sd   = sd(P_Acc)   %>% round(2)
  )
```

Responses that did not use a pronoun are infrequent and evenly distributed across pronoun conditions.

```{r}
#| label: exp1-production-pronoun-dist

table(exp1a_d$Pronoun, exp1a_d$P_Pronoun)

(exp1a_d %>% filter(P_Pronoun == "none") %>% pull(P_Pronoun) %>% length()) /
(exp1a_d %>% pull(P_Pronoun) %>% length())
```
```{r}
#| label: exp1-production-pronoun-errors

exp1a_d %>%
  filter(P_Pronoun != "none") %>%
  group_by(Pronoun, P_Pronoun) %>%
  summarise(n = n()) %>%
  mutate(prop = (n / (n_distinct(exp1a_d$SubjID) * 4)) %>% round(2))
```

60% of participants produced they/them at least once.

```{r}
#| label: exp1-production-usedThey

exp1a_r_prod_usedThey <- exp1a_d %>%
  filter(P_Pronoun == "they/them") %>%
  summarize(n = n_distinct(SubjID))

exp1a_r_prod_usedThey / (n_distinct(exp1a_d$SubjID))
```

## Model

Same model specifications as first model (memory accuracy). Here, the maximal model has random intercepts and slopes by participant, and no random effects by item.

```{r}
#| label: exp1-production-model
#| cache: true

exp1a_m_prod <- buildmer(
  formula = P_Acc ~ Pronoun + (1 + Pronoun | SubjID) + (1 + Pronoun | Name),
  data = exp1a_d, family = binomial,
  buildmerControl(direction = "order")
)

summary(exp1a_m_prod)
```

Convert to odds:

```{r}
#| label: exp1-production-OR

# intercept (mean)
exp1a_m_prod@model %>% get_intercept() %>% exp()

# they/them vs. he/him + she/her
exp1a_m_prod@model %>%
  get_parameters() %>%
  filter(Parameter == "Pronounthey vs he+she") %>%
  pull(Estimate) %>%
  exp()
```

-   The intercept is significant (p\<.001), such that participants are 3.78 times more likely to answer correctly than incorrectly across all pronoun types.

-   The contrast between they/them and he/him + she/her is significant (p\<.001), such that participants are 62.92 times more likely to get he/him and she/her right than they/them.

-   The contrast between he/him and she/her is not significant.

# Memory Predicting Production

## Descriptive Stats

Accuracy for producing they/them is lower than accuracy for remembering they/them. But for he/him and she/her, production accuracy is higher.

```{r}
#| label: exp1-compare-means

exp1a_d %>%
  pivot_longer(
    cols      = c(M_Acc, P_Acc),
    names_to  = "Task",
    values_to = "Acc"
  ) %>%
  group_by(Pronoun, Task) %>%
  summarise(mean = mean(Acc) %>% round(2))
```

Combining the two measures, there are 4 possible patterns: getting both right, getting both wrong, getting just memory right, and getting just production right.

```{r}
#| label: exp1-memory-production-dist

exp1a_r_dist <- exp1a_d %>%
  mutate(Combined_Accuracy = case_when(
    M_Acc == 1 & P_Acc == 1 ~ "Both right",
    M_Acc == 0 & P_Acc == 0 ~ "Both wrong",
    M_Acc == 1 & P_Acc == 0 ~ "Memory only",
    M_Acc == 0 & P_Acc == 1 ~ "Production only"
  )) %>%
  group_by(Pronoun, Combined_Accuracy) %>%
  summarise(n = n())

exp1a_r_dist
```

Production accuracy for they/them when memory was correct vs incorrect.

```{r}
#| label: exp1-memory-production-split

exp1a_d %>%
  filter(Pronoun == "they/them") %>%
  group_by(M_Acc, Pronoun) %>%
  summarise(P_Acc = mean(P_Acc) %>% round(2))
```

## Model

```{r}
#| label: exp1-mean-center-acc

exp1a_d %<>% mutate(M_Acc_Factor = as.factor(M_Acc))
contrasts(exp1a_d$M_Acc_Factor) <- cbind("wrong vs right" = c(-0.5, +0.5))
contrasts(exp1a_d$M_Acc_Factor)
```

Model predicting production accuracy with pronoun type and memory accuracy. Otherwise the same model specifications as the first two. The maximal model includes random intercepts by participant.

```{r}
#| label: exp1-memory-production-model
#| cache: true

exp1a_m_mp <- buildmer(
  formula = P_Acc ~ M_Acc_Factor * Pronoun +
    (Pronoun | SubjID) + (Pronoun | Name),
  data = exp1a_d, family = binomial,
  buildmerControl(direction = "order")
)

summary(exp1a_m_mp)
```

Convert to odds:

```{r}
#| label: exp1-memory-production-OR

# memory accuracy
exp1a_m_mp@model %>%
  get_parameters() %>%
  filter(Parameter == "M_Acc_Factorwrong vs right") %>%
  pull(Estimate) %>%
  exp()

# they/them vs. he/him + she/her * memory accuracy
exp1a_m_mp@model %>%
  get_parameters() %>%
  filter(Parameter == "Pronounthey vs he+she:M_Acc_Factorwrong vs right") %>%
  pull(Estimate) %>%
  exp()
```

-   The effect of memory accuracy is significant (p\<.001), such that participants are 3.47x more likely to get the production right if they got the memory right.

-   Significant interaction between pronoun type (they/them vs. he/him + she/her) and memory accuracy (p\<.05) (odds 0.44). The relative difficulty of they/them was attenuated when the participant had correctly remembered the character's pronoun during the memory phase of the task.

# Compare Memory and Production

Compare accuracy between memory and production tasks by mean-center effects coding Task:

```{r}
#| label: exp1-task-setup

exp1a_d_task <- exp1a_d_all %>%  # memory + prod in long format
  filter(Task != "introduction") %>%
  filter(M_Type == "pronoun" | M_Type == "") %>%
  select(Task, Pronoun, Name, M_Acc, P_Acc, SubjID, Name) %>%
  mutate(.keep = c("unused"), Acc = case_when(
    !is.na(M_Acc) ~ M_Acc,  # group into 1 accuracy variable
    !is.na(P_Acc) ~ P_Acc
  )) %>%
  mutate(  # add pronoun dummy code variables
    Pronoun_They0  = ifelse(Pronoun == "they/them", 0, 1),
    Pronoun_HeShe0 = ifelse(Pronoun != "they/them", 0, 1)
  )

# Mean-center effects code Task
exp1a_d_task$Task %<>% factor()
contrasts(exp1a_d_task$Task) <- cbind(
  "mem vs prod" = c(-.5, +.5)
)
contrasts(exp1a_d_task$Task)
```

```{r}
#| label: exp1-task-model-all
#| cache: true

exp1a_m_task_all <- buildmer(
  formula = Acc ~ Pronoun * Task +
    (Pronoun | SubjID) + (Pronoun | Name),
  data = exp1a_d_task,
  family = binomial,
  buildmerControl(direction = "order")
)

summary(exp1a_m_task_all)
```

-   No main effect of Task (p = .22)
-   Significant interaction between Pronoun and Task (beta = 1.21, p <. 001)

Dummy code to probe interaction:

First, the Task effect just for they/them characters:

```{r}
#| label: exp1-task-model-they
#| cache: true

exp1a_m_task_they0 <- glmer(
  formula = Acc ~ Pronoun_They0 * Task +
    (1 | SubjID) + (1 | Name),  # use random effects from main
  data = exp1a_d_task,
  family = binomial,
)

summary(exp1a_m_task_they0)
```

-   Production is significantly *less* accurate than memory for they/them characters (beta = -0.81, p <. .001)

Second, the Task effect just for he/him and she/her characters:

```{r}
#| label: exp1-task-model-heshe
#| cache: true

exp1a_m_task_heshe0 <- glmer(
  formula = Acc ~ Pronoun_HeShe0 * Task +
    (1 | SubjID) + (1 | Name),  # use random effects from main
  data = exp1a_d_task,
  family = binomial
)

summary(exp1a_m_task_heshe0)
```

- Production is significantly *more* accurate than memory for he.him and she/her characters (beta = 0.55, p <. .001)    

# Compare Pet Questions

Compare pet accuracy to pronoun accuracy. Pronoun (renamed to Character Pronoun here for clarity) stays contrast coded as before, and Question Type (M_Type = pet or pronoun) is mean-center effects coded, comparing pet questions to pronoun questions.

```{r}
#| label: exp1-compare-all-means

exp1a_d_all %>%
  group_by(M_Type, Pronoun) %>%
  filter(!is.na(M_Acc)) %>%  # production rows
  summarise(
    mean = mean(M_Acc) %>% round(2),
    sd   = sd(M_Acc)   %>% round(2)
  )

exp1a_d_pronounsPets <- exp1a_d_all %>%
  filter(M_Type == "pet" | M_Type == "pronoun") %>%
  rename("CharPronoun" = "Pronoun")

exp1a_d_pronounsPets$M_Type %<>% droplevels()
contrasts(exp1a_d_pronounsPets$M_Type) <- cbind("petQ vs pronounQ" = c(-.5, .5))
contrasts(exp1a_d_pronounsPets$M_Type)
```

```{r}
#| label: exp1-compare-pets-model
#| cache: true

exp1a_m_pet <- buildmer(
  formula = M_Acc ~ CharPronoun * M_Type +
    (M_Type * CharPronoun | SubjID) +
    (M_Type * CharPronoun | Name),
  data = exp1a_d_pronounsPets, family = binomial,
  buildmerControl(direction = "order")
  )

summary(exp1a_m_pet)
```

-   Significant main effect of Question Type (p\<.001), with higher accuracy for pronouns (this makes sense seeing the high accuracy rates for he/him and she/her pronoun questions).
-   Significant interaction between Pronoun and Question Type, such that the difference between Question Types is larger for he/him + she/her questions than for they/them characters.

To check this interaction direction, dummy coded Pronoun with they/them characters as 0 and he/him and she/her characters as 1:

```{r}
#| label: exp1-compare-pets-model-they0
#| cache: true

exp1a_d_pronounsPets %<>% mutate(CharPronoun_They0 = ifelse(
  CharPronoun == "they/them", 0, 1))

exp1a_m_pet_they <- glmer(
  formula = M_Acc ~ CharPronoun_They0 * M_Type +
    (M_Type | SubjID),  # use random effects from main
  data = exp1a_d_pronounsPets,
  family = binomial
)

summary(exp1a_m_pet_they)
```

-   Non-significant main effect of Question Type (M_Type) means that there is no difference between pet and pronoun questions when Pronoun is 0 (= for they/them characters).

Now dummy code to get main effect of Question Type in he/him + she/her (= 0)

```{r}
#| label: exp1-compare-pets-model-heshe0
#| cache: true

exp1a_d_pronounsPets %<>% mutate(CharPronoun_HeShe0 = ifelse(
  CharPronoun == "they/them", 1, 0))

exp1a_m_pet_heshe <- glmer(
  formula = M_Acc ~ CharPronoun_HeShe0 * M_Type +
      (M_Type | SubjID),  # use random effects from main
  data = exp1a_d_pronounsPets,
  family = binomial
)

summary(exp1a_m_pet_heshe)
```

-   Significant (p \< .001) main effect of Question Type (M_Type) means that the difference is in he/him + she/her characters (but not they/them)
