---
title: "Improving memory for and production of singular <i>they</i> pronouns: Experiment 2"
author: "Bethany Gardner"
date: "04/13/2023"
output: 
  github_document:
    toc: true
    toc_depth: 3
editor_options: 
  markdown: 
    wrap: 72
---

```{r}
#| label: exp2-setup
#| include: false

library(tidyverse)
library(magrittr)
library(broom.mixed)
library(lmerTest)
library(buildmer)
library(sjmisc)

options(dplyr.summarise.inform = FALSE)
```

# Load data

Read data, preprocessed from PCIbex output. See `data/exp2_data_readme` for more details.

```{r}
#| label: exp2-load

exp2_d_all <- read.csv("data/exp2_data.csv", stringsAsFactors = TRUE) %>%
  rename("Biographies" = "Story") # rename to match labeling in paper

exp2_d_all$Participant %<>% as.factor()

exp2_d_all$PSA %<>% as.factor() %>%
  recode_factor("0" = "Unrelated", "1" = "Gender")

exp2_d_all$Biographies %<>% as.factor() %>%
  recode_factor("0" = "HeShe", "1" = "They")

exp2_d_all$X <- NULL

str(exp2_d_all)
```

Set up contrast coding for Pronoun Type. The first contrast compares *they* to *he*+*she*. The second contrast compares *he* to *she*.

```{r}
#| label: exp2-contrasts-1

contrasts(exp2_d_all$Pronoun) <- cbind(
  "_T_HS" = c(.33, .33, -.66),
  "_H_S"  = c(-.5, .5, 0)
)
contrasts(exp2_d_all$Pronoun)
```

Set up contrast coding for PSA and Biographies conditions. .5 are the conditions related to singular they (gendered language PSA, they/them biographies); -.5 are the unrelated conditions (unrelated PSA, he/him and she/her biographies).

```{r}
#| label: exp2-contrasts-2

# check labels
exp2_d_all %>% count(Condition, PSA, Biographies)

contrasts(exp2_d_all$PSA) <- cbind("_GenLang" = c(-.5, .5))
contrasts(exp2_d_all$PSA)

contrasts(exp2_d_all$Biographies) <- cbind("_They" = c(-.5, .5))
contrasts(exp2_d_all$Biographies)
```

Remove pet and job rows, and the columns that aren't used in the models.

```{r}
#| label: exp2-data-subset

exp2_d <- exp2_d_all %>%
  filter(M_Type == "pronoun") %>%
  select(
    Participant, Condition, PSA, Biographies, Name, Pronoun,
    M_Acc, M_Response, P_Acc, P_Response
  )

str(exp2_d)
summary(exp2_d)
```

# Memory

## Descriptive Stats

Mean accuracy for all three memory question types.

```{r}
#| label: exp2-memory-all-means

exp2_d_all %>%
  group_by(M_Type) %>%
  summarise(
    mean = mean(M_Acc) %>% round(2),
    sd   = sd(M_Acc)   %>% round(2)
  )
```

Mean accuracy, split by Pronoun Type, PSA, and Biographies conditions. [Both = gendered language PSA + *they* biographies; PSA = gendered language PSA + *he*/she biographies; Story = unrelated PSA + they biographies; Neither = unrelated PSA + he/she biographies.]

```{r}
#| label: exp2-memory-condition-means

exp2_d %>%
  mutate(Pronoun_Group = ifelse(Pronoun == "they/them", "They", "HeShe")) %>%
  group_by(Pronoun_Group, PSA, Biographies) %>%
  summarise(mean = mean(M_Acc) %>% round(2)) %>%
  pivot_wider(names_from = Pronoun_Group, values_from = mean) %>%
  mutate(Mean_Diff = HeShe - They) %>%
  arrange(Mean_Diff)
```

90-95% of participants selected they/them at least once.

```{r}
#| label: exp2-memory-usedThey

exp2_d %>%
  filter(M_Response == "they/them") %>%
  group_by(Condition) %>%
  summarise(
    n    = n_distinct(Participant),
    prop = n / 80
  )
```

## Model

Full model has interactions between Pronoun (2 contrasts), PSA, and Biographies; random intercepts and slopes by participant and item. buildmer finds the maximal model that will converge (but doesn't then go backward to remove non-significant terms, the default setting). The final model includes all fixed effects/interactions and random intercepts by name.

```{r}
#| label: exp2-memory-model
#| cache: true

exp2_m_memory <- buildmer(
  formula = M_Acc ~ Pronoun * PSA * Biographies +
    (Pronoun | Participant) + (Pronoun | Name),
  data = exp2_d, family = binomial,
  buildmerControl(direction = "order")
)
summary(exp2_m_memory)
```

# Production

## Descriptive Stats

Mean accuracy, split by Pronoun Type, PSA, and Biographies conditions. [Both = gendered language PSA + *they* biographies; PSA = gendered language PSA + *he/she* biographies; Story = unrelated PSA + *they* biographies; Neither = unrelated PSA + *he/she* biographies.]

```{r}
#| label: exp2-production-means

exp2_d %>%
  mutate(Pronoun_Group = ifelse(Pronoun == "they/them", "They", "HeShe")) %>%
  group_by(Pronoun_Group, PSA, Biographies) %>%
  summarise(mean = mean(P_Acc) %>% round(2)) %>%
  pivot_wider(names_from = Pronoun_Group, values_from = mean) %>%
  mutate(Mean_Diff = HeShe - They) %>%
  arrange(Mean_Diff)
```

## Model

Same model specifications as before. The maximal model contains all fixed effects/interactions and by-item random intercepts.

```{r}
#| label: exp2-production-model
#| cache: true

exp2_m_prod <- buildmer(
  formula = P_Acc ~ Pronoun * PSA * Biographies +
    (Pronoun | Participant) + (Pronoun | Name),
  data = exp2_d, family = binomial,
  buildmerControl(direction = "order")
)
summary(exp2_m_prod)
```

## Three-Way Interaction

The main model has Helmert coding for Pronoun and Effects coding (.5, -.5) for PSA and Biographies. This means Pronoun (T vs HS) \* PSA \* Biographies is testing the interaction between Pronoun and PSA across both Biographies conditions.

Dummy coding Biographies with *they* biographies as 1 and *he/she* biographies as 0 tests the interaction between Pronoun and PSA for just the *he/she* Biographies:

```{r}
#| label: exp2-production-interaction-HS-bios
#| cache: true

exp2_d %<>% mutate(Biographies_HS0 = Biographies)
contrasts(exp2_d$Biographies_HS0) <- cbind("_T1" = c(0, 1))
contrasts(exp2_d$Biographies_HS0)

exp2_m_prod_bio_HS0 <- glmer(
  formula = P_Acc ~ Pronoun * PSA * Biographies_HS0 + (1 | Name),
  data = exp2_d, family = binomial
)
summary(exp2_m_prod_bio_HS0)
```

Conversely, dummy coding Biographies with *he/she* biographies as 1 and *they* biographies as 0 tests the interaction between Pronoun and PSA for just the *they* Biographies.

```{r}
#| label: exp2-production-interaction-T0
#| cache: true

exp2_d %<>% mutate(Biographies_T0 = Biographies)
contrasts(exp2_d$Biographies_T0) <- cbind("_HS1" = c(1, 0))
contrasts(exp2_d$Biographies_T0)

exp2_m_prod_bio_T0 <- glmer(
  formula = P_Acc ~ Pronoun * PSA * Biographies_T0 + (1 | Name),
  data = exp2_d, family = binomial
)
summary(exp2_m_prod_bio_T0)
```

The three models to compare:

```{r}
#| label: exp2-production-interaction-results

exp2_r_prod_interaction <- bind_rows(
    .id = "model",
    "Across_Bio" = exp2_m_prod@model   %>% tidy(),
    "HeShe_Bio"  = exp2_m_prod_bio_HS0 %>% tidy(),
    "They_Bio"   = exp2_m_prod_bio_T0  %>% tidy()
  ) %>%
  select(model, term, estimate, p.value) %>%
  filter(
    term == "Pronoun_T_HS" |
    term == "PSA_GenLang" |
    term == "Pronoun_T_HS:PSA_GenLang"
  ) %>%
  mutate(
    estimate = round(estimate, 4),
    p.value  = round(p.value, 4)
  ) %>%
  arrange(term, model)

exp2_r_prod_interaction
```

The estimate for the PSA\*Pronoun interaction is -2.34 for the *he/she* biographies and -1.47 for the *they* biographies, which means that the pronoun PSA reduced the relative difficulty of they/them more when paired with the *he/she* biographies than with *they* biographies. Connecting to the barplot, the PSA-Neither difference is larger than Both-Story difference.

## Producing they/them at least once

```{r}
#| label: exp2-produced-they-means

exp2_d %>%
  filter(P_Response == "they/them") %>%
  group_by(Condition) %>%
  summarise(
    n    = n_distinct(Participant),
    prop = n / 80
  )
```

Model with whether each participant produced they/them at least once as the outcome variable. Higher with the gendered language PSA, no effect of Biographies, vaguely trending interaction.

```{r}
#| label: exp2-produced-they-model

exp2_d_they <- exp2_d %>%
  mutate(
    M_IsThey = ifelse(M_Response == "they/them", 1, 0),
    P_IsThey = ifelse(P_Response == "they/them", 1, 0)
  ) %>%
  group_by(Participant, Condition, PSA, Biographies) %>%
  summarise(
    M_Count = sum(M_IsThey),
    P_Count = sum(P_IsThey)
  ) %>%
  mutate(
    M_UseThey = ifelse(M_Count != 0, 1, 0),
    P_UseThey = ifelse(P_Count != 0, 1, 0)
  )

exp2_m_prod_useThey <- glm(
  formula = P_UseThey ~ PSA * Biographies,
  data = exp2_d_they, family = binomial
)
summary(exp2_m_prod_useThey)
```

# Memory Predicting Production

## Descriptive Stats

```{r}
#| label: exp2-memory-production-means

exp2_d %>%
  group_by(Pronoun, PSA, Biographies, M_Acc) %>%
  summarise(m = mean(P_Acc))
```

Combining the two measures, there are 4 possible patterns: getting both right, getting both wrong, getting just memory right, and getting just production right.

```{r}
#| label: exp2-memory-production-dist

exp2_r_dist <- exp2_d %>%
  mutate(
    Combined_Accuracy = case_when(
      M_Acc == 1 & P_Acc == 1 ~ "Both right",
      M_Acc == 0 & P_Acc == 0 ~ "Both wrong",
      M_Acc == 1 & P_Acc == 0 ~ "Memory only",
      M_Acc == 0 & P_Acc == 1 ~ "Production only"
    )
  ) %>%
  group_by(Pronoun, PSA, Biographies, Combined_Accuracy) %>%
  summarise(n = n())

exp2_r_dist
```

Production accuracy for they/them when memory was correct vs incorrect.

```{r}
#| label: exp2-memory-production-split

exp2_d %>%
  filter(Pronoun == "they/them") %>%
  group_by(Condition, M_Acc) %>%
  summarise(P_Acc = mean(P_Acc))
```

## Model

```{r}
#| label: exp2-memory-factor

exp2_d %<>% mutate(M_Acc_Factor = as.factor(M_Acc))

contrasts(exp2_d$M_Acc_Factor) <- cbind("_Wrong_Right" = c(-0.5, +0.5))
contrasts(exp2_d$M_Acc_Factor)
```

Maximal model has interactions between Pronoun (2 contrasts), Memory Accuracy, PSA, and Biographies, then random intercepts by item.

```{r}
#| label: exp2-memory-production-model
#| cache: true

exp2_m_mp <- buildmer(
  formula = P_Acc ~ Pronoun * PSA * Biographies * M_Acc_Factor +
    (Pronoun | Participant) + (Pronoun | Name),
  data = exp2_d, family = binomial,
  buildmerControl(direction = "order")
)
summary(exp2_m_mp)
```

# Compare Pet Questions

Compare pet accuracy to pronoun accuracy. Pronoun (renamed to Character Pronoun here for clarity) stays contrast coded as before and Question Type (M_Type = pet or pronoun) is mean-center effects coded, comparing pet questions to pronoun questions.

```{r}
#| label: exp2-compare-pets-means

# mean and sd for pets
exp2_d_all %>%
  filter(M_Type == "pet") %>%
  pull(M_Acc) %>%
  mean()
exp2_d_all %>%
  filter(M_Type == "pet") %>%
  pull(M_Acc) %>%
  sd()

# subset data
exp2_d_pronounsPets <- exp2_d_all %>%
  filter(M_Type == "pet" | M_Type == "pronoun") %>%
  rename("CharPronoun" = "Pronoun")

# check other contrasts are still in df
contrasts(exp2_d_pronounsPets$CharPronoun)
contrasts(exp2_d_pronounsPets$PSA)
contrasts(exp2_d_pronounsPets$Biographies)

# mean-center effects code question type
exp2_d_pronounsPets$M_Type %<>% droplevels()
contrasts(exp2_d_pronounsPets$M_Type) <- cbind(
  "=Pet_Pronoun" = c(-.5, .5)
)
contrasts(exp2_d_pronounsPets$M_Type)
```

```{r}
#| label: exp2-compare-pets-model
#| cache: true

exp2_m_pet <- buildmer(
  formula = M_Acc ~ CharPronoun * PSA * Biographies + # conditions
    M_Type + # add question type
    CharPronoun * M_Type + # but only its interaction with Pronoun
    (M_Type * CharPronoun | Participant) +
    (M_Type * CharPronoun | Name),
  data = exp2_d_pronounsPets, family = binomial,
  buildmerControl(direction = "order")
)
summary(exp2_m_pet)
```

```{r}
#| label: exp2-compare-pets-model-they0
#| cache: true

exp2_d_pronounsPets %<>% mutate(CharPronoun_They0 = ifelse(
  CharPronoun == "they/them", 0, 1
))

exp2_m_pet_they <- glmer( # same as buildmer results, just swap CharPronoun
  formula = M_Acc ~ M_Type + CharPronoun_They0 + M_Type:CharPronoun_They0 +
    PSA + Biographies + CharPronoun_They0:PSA + PSA:Biographies +
    CharPronoun_They0:Biographies +
    CharPronoun_They0:PSA:Biographies +
    (1 + M_Type | Name) + (1 | Participant),
  data = exp2_d_pronounsPets, family = binomial
)
summary(exp2_m_pet_they)
```

```{r}
#| label: exp2-compare-pets-model-heshe
#| cache: true

exp2_d_pronounsPets %<>% mutate(CharPronoun_HeShe0 = ifelse(
  CharPronoun == "they/them", 1, 0
))

exp2_m_pet_heshe0 <- glmer( # same as buildmer results, just swap CharPronoun
  formula = M_Acc ~ M_Type + CharPronoun_HeShe0 + M_Type:CharPronoun_HeShe0 +
    PSA + Biographies + CharPronoun_HeShe0:PSA + PSA:Biographies +
    CharPronoun_HeShe0:Biographies +
    CharPronoun_HeShe0:PSA:Biographies +
    (1 + M_Type | Name) + (1 | Participant),
  data = exp2_d_pronounsPets, family = binomial
)
summary(exp2_m_pet_heshe0)
```

# Compare to Exp1

```{r}
#| label: compare-exp1-exp2-memory

bind_rows(
  read.csv("data/exp1a_data.csv", stringsAsFactors = TRUE) %>%
    filter(M_Type == "pronoun") %>%
    group_by(Pronoun) %>%
    summarise(
      mean = mean(M_Acc) %>% round(2),
      sd   = sd(M_Acc)   %>% round(2),
      exp  = 1
    ),
  exp2_d %>%
    filter(PSA == 0 & Biographies == 0) %>%
    group_by(Pronoun) %>%
    summarise(
      mean = mean(M_Acc) %>% round(2),
      sd   = sd(M_Acc)   %>% round(2),
      exp  = 2
    )
)
```

```{r}
#| label: compare-exp1-exp2-production

bind_rows(
  read.csv("data/exp1a_data.csv", stringsAsFactors = TRUE) %>%
    filter(!is.na(P_Acc)) %>%
    group_by(Pronoun) %>%
    summarise(
      mean = mean(P_Acc) %>% round(2),
      sd   = sd(P_Acc)   %>% round(2),
      exp  = 1
    ),
  exp2_d %>%
    filter(PSA == 0 & Biographies == 0) %>%
    group_by(Pronoun) %>%
    summarise(
      mean = mean(P_Acc) %>% round(2),
      sd   = sd(P_Acc)   %>% round(2),
      exp  = 2
    )
  )
```

Accuracy for unrelated job question:

```{r}
#| label: compare-exp1-exp2-job

bind_rows(
  read.csv("data/exp1a_data.csv", stringsAsFactors = TRUE) %>%
    filter(M_Type == "job") %>%
    summarise(
      mean = mean(M_Acc) %>% round(2),
      sd   = sd(M_Acc)   %>% round(2),
      exp  = 1
    ),
  exp2_d_all %>%
    filter(M_Type == "job") %>%
    filter(PSA == 0 & Biographies == 0) %>%
    summarise(
      mean = mean(M_Acc) %>% round(2),
      sd   = sd(M_Acc)   %>% round(2),
      exp  = 2
    )
  )
```

And for comparison pet question:

```{r}
#| label: compare-exp1-exp2-pet

bind_rows(
  read.csv("data/exp1a_data.csv", stringsAsFactors = TRUE) %>%
    filter(M_Type == "pet") %>%
    summarise(
      mean = mean(M_Acc) %>% round(2),
      sd   = sd(M_Acc) %>% round(2),
      exp  = 1
    ),
  exp2_d_all %>%
    filter(M_Type == "pet") %>%
    filter(PSA == "Gender" & Biographies == "They") %>%
    summarise(
      mean = mean(M_Acc) %>% round(2),
      sd   = sd(M_Acc) %>% round(2),
      exp  = 2
    )
  )
```
